---
title: "Shiny Demo"
output: ioslides_presentation
---


## What is Shiny?
- Framework for developing web apps in R
- Does not require knowledge of traditional web technologies (HTML, CSS, JavaScript)
- Automatically calculates which parts of the app to update when inputs change (reactive programming)

## Contents
- Basic app structure
- UI components: inputs, outputs and layouts
- Reactive programming
- Reactive expressions
- The reactive graph
- Modules
- eventReactive and observeEvent

## Setup
- To follow along with this demo, you can clone or download the GitHub repository at:
- https://github.com/YipengC/shiny_demo

## Default example app
- `src/01_default` contains the default app created when a Shiny app project is 
initialised in RStudio
- This can be started by opening `app.R` and clicking `Run App` in RStudio,
- or calling `shiny::runApp("src/01_default")` from the console
- You can also run the app by calling `source()` on the `app.R` file if you're
running from outside of RStudio
- `shiny::runApp("src/01_default", display.mode = "showcase")` shows the corresponding
code while the app is running

## Shiny app structure
```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(

)

server <- function(input, output, session) {

}

# Run the application 
shinyApp(ui = ui, server = server)

```
- ui: defines the components (layouts, inputs, outputs) and how they are arranged
- server: specifies how these components and data interact

# UI components

## Inputs
- Defined in the ui using a function call (ending with `Input()`)
- Parameters include the `inputId`, `label` and other input-specific variables
- e.g.
```{r, eval=FALSE}
sliderInput(inputId = "bins",
            label = "Number of bins:",
            min = 1,
            max = 50,
            value = 30)
```
- Some example inputs are shown in `src/input_examples`

## Outputs
- Creates a placeholder to be filled by the server function
- Defined in ui function using a function (ending with `Output()`)
- Parameters include the `outputId` and other output-specific variables
- e.g.
```{r eval=FALSE}
plotOutput(outputId = "distPlot")
```
- Some example outputs are shown in `src/output_examples`

## Layouts
- Used to define the arrangement of inputs and outputs in the UI
- Defined by a hierarchy of function calls in the ui function
```{r, eval=FALSE}
ui <- fluidPage(

    # Application title
    titlePanel(),
    sidebarLayout(
        sidebarPanel(

        ),
        mainPanel(

        )
    )
)
```

# Reactive programming

## What is reactive programming?
- A programming paradigm
- Specify a graph of dependencies
- Outputs are automatically computed when an input is changed

## The server function: input
```{r, eval=FALSE}
server <- function(input, output, session) {}
```
- `input`: a named list of input data sent from the browser. The names correspond 
to the inputIds specified in the ui function
- `input` is read-only
- `input` can only be read within a *reactive context*

## The server function: output
```{r, eval=FALSE}
server <- function(input, output, session) {}
```
- `output`: a named list of outputs specified in the ui function. The names correspond
to the outputIds.
- Used to send output to the UI
- outputs are used with a `render` function, which creates a *reactive context* 
which keeps track of which inputs it uses
- The `render` function also converts the result into a format the browser can display

## A closer look at using an output
```{r, eval=FALSE}
output$distPlot <- renderPlot({
    # generate bins based on input$bins from ui.R
    x    <- faithful[, 2]
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    # draw the histogram with the specified number of bins
    hist(x, breaks = bins, col = 'darkgray', border = 'white',
         xlab = 'Waiting time to next eruption (in mins)',
         main = 'Histogram of waiting times')
})
```
- The code within the `{}` is not executed immediately when `renderPlot` is called.
Instead it is executed whenever the output `distPlot` needs to update.
- With the `renderPlot` function, we specify how the output `distPlot` should be 
computed, whenever its dependencies (input$bins) change. Note: this affects execution order


# Example app

## London Marathon winner dataset
- For the remaining examples, we'll use the TidyTuesday London Marathon winners dataset
```{r}
df <- readRDS("../data/marathon_winners.rds")
head(df)
```

## 03_basic_app
- Plot a histogram of times given a filter for the years and category

## 04_two_plots
- Create a second plot showing time vs. year

## 05_two_plots_filtered_data_reactive
- Currently we are performing the filter in each of the `render` functions
- This can be made more efficient by using a *reactive expression*
- Reactive expressions are another way of creating a *reactive context*

## The reactive graph
- Now that we know about inputs, outputs and reactive expressions, we'll look at
how the reactive graph determines how dependencies are executed and resolved.
- Some terminology
  - producers and consumers
  - inputs, reactive expressions and outputs
- https://mastering-shiny.org/reactive-graph.html
- Shiny reactlog

## Modules 06/07
- Manage complexity
- Abstract out reusable components
- Modules can be tested in isolation
- Namespaces/managing IDs

## 08_update_on_button_click
- Used for more traditional event handling e.g. on clicking a button
- `eventReactive(eventExpr, handlerExpr)`: for calculating values - returns a reactive
- `observeEvent(eventExpr, handlerExpr)`: for performing actions with side effects - does not return a reactive
- These functions will execute `handlerExpr` only when `eventExpr` changes and does not 
introduce a dependency on any other reactives in `handlerExpr` (unlike `reactive()`)